---
title: "Project Team Six"
subtitle: "Proposal"
format: html
editor: visual
---

```{r}
#| label: load-pkgs
#| message: false

library(tidyverse)
```

# Data 1

## Introduction and data

This dataset is aggregated from the FBI Hate Crimes Report for 2013. The data was collected in 2016 by accessing statistics across states from the FBI Hate Crimes Report website, which is open access (<https://www.fbi.gov/about-us/cjis/ucr/hate-crime/2013>). The dataset contains two csv files with a table detailing demographics for agencies that reported incidences of hate crimes in 2013 and another for those that reported no incidences of hate crimes in 2013. These tables contain 1826 and 12826 observations, respectively. The first table contains the variables state, agency type (city, university, county, etc.), agency name, and the number of occurrences in each motivation category of hate crime --- race, religion, sexual orientation, ethnicity, disability, gender, and gender identity. Additionally, there are columns that indicate the number of hate crime occurrences in each quarter. Finally, the last column details the population of the location of the hate crime. The second table, because it omits the motivation details, only contains columns for state, agency type and name, quarter, and population. There are a few ethical concerns with this data. Perhaps most prominent, as the literature review later mentions, the incidents in this report are only those that law enforcement has determined meets the criteria for a hate crime. This means, if such criteria is especially constrictive, the dataset may be underreporting the true frequency of hate crimes. 

## Research question

**How has hate crime victimization and frequency changed since 2013?**

-   This question is important because it can be used to quantify social progress in our country, as well as inform stats-based policy reform and resource allocation for minority populations.

-   In an ethnically, racially, religiously and sexually diverse nation, hate crimes are one of the many substantial barriers to our functioning as a safe and accepting society. 2013 was a year in which the political climate of the United States was less accepting of diverse identities compared to today's social atmosphere. Specifically, 2013 was pre- legalization of same sex marriage and pre- black lives matter. I hypothesize that the data (when analyzed along with more recent data) will reveal a decrease in hate crimes targeting persons with disabilities as well as those targeting religious, sexuality, and gender minorities since 2013, with an increase in hate crimes targeting racial and ethnic minorities. Overall, I predict higher levels of hate crimes in the race, ethnicity, sexual orientation, gender and gender identity categories. I also think that there will be higher occurrences of hate crimes in general in historically red states/counties.

-   The variables state and agency type are both categorical. The population variable and variables for the number of hate crimes in each quarter and minority category are quantitative.

## Literature

<https://doi.org/10.1037/sgd0000227>

This source discusses the documentation and data reporting of hate crimes in the United States. Specifically, it details which of the most commonly cited sources are the most reliable. While this article is looking at only hate crimes targeting gender and sexual minorities, I think the information is useful and relevant regardless. The paper concludes that reports made to local (non-law enforcement) organizations are valuable sources of data, but this can exclude a number of people such as closeted individuals and those living in areas that are not served by such an organization. As for FBI data, the strengths lie in the reliability, accuracy, and accessibility, as well as the guarantee that the observations included are only those confirmed to meet the standardized legal criteria of a hate crime. However, the aforementioned strength can also be a weakness. Because many hate crimes are not fully investigated and confirmed to be hate crimes, or not reported by victims at all, the data is likely reporting a significant underestimate of the true frequency of these crimes.

The literature above is useful for determining which sources are the most accurate for reporting on hate crime frequency, and it specifically details the pros and cons of using the FBI's hate crime report data. Our project can utilize this literature by acknowledging those strengths and weaknesses in our own research, and possibly looking for other reliable sources reporting on the same years to compare with the FBI data.

## Glimpse of data

```{r}
#| label: load-data-1

hatecrime <- read_csv("data/table13.csv")
nohate <- read_csv("data/table14.csv")

glimpse(hatecrime)
glimpse(nohate)
```

# Data 2

## Introduction and data

-   Identify the source of the data.

The source of the data comes from this website: https://data.world/datadavis/nba-salaries#.

-   State when and how it was originally collected (by the original data curator, not necessarily how you found the data).

The data originated from Basketball Reference. The data was collected during the 2017-2018 NBA season.

-   Write a brief description of the observations.

Each observation represents an NBA player in a specific NBA season. There are 30 variables that consist of the player ID, the salary they made in a particular season when the season began and ended, the team the NBA player played for, the name of the NBA player, the birthplace and birthdate of the NBA player, the high school and college the NBA player played for, the year, pick, and team the player was drafted for, the NBA career statistics for the NBA player, the height and weight of the NBA player, the position the NBA player plays, and the NBA player's preferred shooting hand. The data is limited for players who played in the NBA between the 1984-1985 NBA season and the 2017-2018 NBA season.

## Research question

-   A well formulated research question. (You may include more than one research question if you want to receive feedback on different ideas for your project. However, one per data set is required.)

Is there a correlation between NBA players' career statistics and their salaries from the 1984-1985 NBA season to the 2017-2018 NBA season?

-   A description of the research topic along with a concise statement of your hypotheses on this topic.

Through this research project, we will examine whether players' career statistics, such as points per game or assists per game directly influence players' salaries. We will employ statistical analysis to uncover any existing correlation in the data set. However, we keep in mind that players' salaries are influenced beyond just statistics, such as the year the player was playing in the NBA and the team the player played for, so we keep that into consideration. We think that several statistics, such as points per game and rebounds per game does affect players' salaries.

-   Identify the types of variables in your research question. Categorical? Quantitative?

Most of the variables we will use are quantitative, such as the player's career statistics (assist per game, points per game, rebounds per game, win shares, field goal percentage, and number of games played) and the player's salary. The categorical variables we will use are the year of the NBA season and possibly the team the player played for.

## Literature

-   Find one published credible article on the topic you are interested in researching.

Stanek, T. (2016). Player performance and team revenues: NBA player salary analysis.

-   Provide a one paragraph summary about the article.

The article conducts a study to determine if NBA players are appropriately paid based on their contributions on the court by looking at and calculating their offensive statistics, defensive statistics, amount of wins the player was responsible for, and age. The article went over the salary cap in the NBA and the CBA (collective bargaining agreement) as it does play a role in how team owners and the NBA distribute player's salaries. The article mentioned that they considered many factors in the player's performance, such as the number of minutes played per game, the position the player plays for, and the amount of wins the player contributed. For example, it doesn't make sense if a player who plays fewer minutes is as valuable as a player who plays more minutes on the court. The article concludes that salaries are affected based on these factors: TV contract revenues, the number of years in the NBA as older players tend to earn more money than younger players, and the amount of points scored per game as players who score more tend to be paid more. However, looking at the stat sheet won't give a full picture of whether or not a player is valuable. Other external factors like injury status, team dynamics, and the player's role in the team need to be considered.

-   In 1-2 sentences, explain how your research question builds on / is different than the article you have cited.

This article builds on what we need to focus on when we consider the relationship between statistics and NBA salaries, such as other factors to consider, like the specific time era the player played in the NBA or if the player plays in a small or large market team.

## Glimpse of data

```{r}
#| label: load-data-2

nba_salaries <- read_csv("data/salaries_1985to2018.csv")
nba_players <- read_csv("data/players.csv")

nba_dataset <- nba_salaries |>
  inner_join(nba_players)

glimpse(nba_dataset)

```

# Data 3

## Introduction and data

-   Identify the source of the data.

    The source of the data comes from this website: 

    <https://www.kaggle.com/datasets/greeshmagirish/crime-against-women-20012014-india>

-   State when and how it was originally collected (by the original data curator, not necessarily how you found the data).

    The data was collected annually from different states and union territories of India from 2001 to 2014 by the National Crime Records Bureau. The Bureau aggregates reports from various police departments across the country.

-   Write a brief description of the observations.

    The dataset contains detailed records of different types of crimes against women, categorized by state and year. The crimes include rape, kidnapping & abduction, dowry deaths, cruelty by husband and relatives, Assault on women with intent to outrage her modesty, Insult to modesty of Women.

## Research question

-   A well formulated research question. (You may include more than one research question if you want to receive feedback on different ideas for your project. However, one per data set is required.)

    How has the rate of specific crimes against women (e.g., dowry deaths, cruelty by husband) evolved over the years in India? Are there specific states where these crimes have seen a significant increase or decrease?

-   A description of the research topic along with a concise statement of your hypotheses on this topic.

    The research will explore the chronological patterns of crimes against women in India, emphasizing variations across different states.The hypothesis is that certain states might show a decline in particular crimes due to proactive interventions, while others could be experiencing increasing numbers, indicating a  need for  measures or intervention.

-   Identify the types of variables in your research question. Categorical? Quantitative?

    -   Crime type : Categorical

    -   Year: Quantitative 

    -   Number of incidents: Quantitative

## Literature

-   Find one published credible article on the topic you are interested in researching.

    "Violence Against Women", National Library of Medicine.

    [**https://www.ncbi.nlm.nih.gov/pmc/articles/PMC4539878/**](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC4539878/)

-   Provide a one paragraph summary about the article.

    This article  delves into the various dimensions of violence against women  in the Indian context. It highlights how the Indian Penal Code  and various local laws classify crimes against women, with specifics such as the increasing rates of rape and state-wise disparities in crime rates. The article  explores the multifaceted consequences of violence on women, highlighting increased chances of re-victimization, depression, suicidal ideation, and health implications. Key causal factors of violence against women are discussed in depth, ranging from socio demographic aspects like poverty and literacy levels, to family factors, psychiatric morbidities, and cultural traditions deeply entrenched in patriarchy. The role of media, addiction, deficiencies in legislation, and an often-inefficient law enforcement machinery also come under scrutiny. The article underscores the interconnectedness of these factors and their collective contribution to the issue, emphasizing the urgent need for holistic solutions and better legislative provisions to address violence against women in India.

-   In 1-2 sentences, explain how your research question builds on / is different than the article you have cited.

    While the cited article provides a comprehensive overview of the types of violence against women in India, its consequences, and root causes, our research specifically focuses on the temporal trends and state-wise variations of these crimes. This enables a more granular understanding of regional disparities and temporal shifts in crimes against women over a defined period.

## Glimpse of data

```{r}
#| label: India Crime Against Women

crime_women <- read_csv("data/crimes_against_women_2001-2014.csv")
glimpse(crime_women)

# add code here
```
